---
title: "Exposure time is an important variable in quantifying post-dispersal seed removal"
author: "Si-Chong Chen, Bo Wang and Angela T. Moles"
date: "16 November 2020"
output: html_document
---

```{r global-options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.width = 7, fig.height = 6)
```

## 1. Exposure time data

The seed removal dataset can be acquired from Dylewski et al. (2020). The dataset deposited in the Dryad Digital Repository has been embargoed till at least 16th November 2020. Through email correspondence, we received the dataset from the authors on 19th October 2020.

```{r include = FALSE}
setwd("S:/Research/MSB/PY and predation/EL")
```

```{r message = FALSE}
# Load packages
library(tidyverse)
library(readxl)
library(glmmTMB)
library(boot)

# Load data
data <- read_excel("DRYAD SEED REMOVAL DATASET.xlsx", sheet = 1, na = "NA")

# Transform data according to Dylewski et al. (2020)
data = data %>% 
  # "Seed mass (+1) was natural log-transformed for analysis."
  mutate(MASS_ln = log(MASS + 1)) %>% 
  # ECOSYSTEM & PROV -- Convert to sentence case
  mutate(ECOSYSTEM = str_to_sentence(ECOSYSTEM ), PROV = str_to_sentence(PROV)) %>% 
  # Reorder factor levels
  mutate(ECOSYSTEM = factor(ECOSYSTEM, levels = c("Grassland", "Temperate", "Tropical", "Other")))
```


We supplemented the original dataset with a column of the exposure time (TIME_D) according to the descriptions in the 71 source publications (PUB).

Note: Three studies (15 data points) did not clearly state the seed exposure times (NA values).
```{r echo = FALSE, message = FALSE}
library(knitr)
library(kableExtra)
data %>% 
  distinct(PUB, TIME_D) %>% 
  arrange(PUB) %>% 
  kable(align = "lr", digits = 2) %>% 
  kable_styling(full_width = F)
```


Distribution of seed predator exposure time by data points (n = 357)

```{r warning = TRUE, fig.width = 6, fig.height = 4}
ggplot(data, aes(TIME_D)) + 
  geom_histogram(binwidth = 1, col=I("black")) +
  xlab("Time (day)") + ylab("Count") +
  theme_classic(base_size = 16)
```

## 2. Reproduce Dylewski et al. (2020) Figure 2

We did not have analyses on ecosystem or provenance, but we followed the figure style as in Dylewski et al. (2020) Figure 2 for the ease of comparison.

```{r}
# Transform data according to Dylewski et al. (2020)
data_C = data %>% 
  # Rescale to 0 < y < 1
  # "... applying a small, linear transformation to eliminate zeros and ones (Smithson & Verkuilen 2006)
  mutate(REMOVAL_C = (REMOVAL * (nrow(data) - 1) + 1/2)/nrow(data))

# Model
# "... using generalised linear mixed models (GLMMs) with a beta distribution and a restricted maximum-likelihood (REML) estimator"
model <- glmmTMB(REMOVAL_C ~ MASS_ln + I(MASS_ln^2) + (1|PUB), data = data_C, 
                 family = beta_family(link = "logit"), REML = TRUE)
summary(model); coef = summary(model)$coefficients$cond


# Palette
cols <- c("Grassland" = "#D55E00", "Temperate" = "#0072B2", "Tropical" = "#FFC425", "Other" = "#888888")

# Graph
ggplot(data = data_C, aes(x = MASS_ln, y = REMOVAL_C)) + 
  geom_point(aes(colour = ECOSYSTEM, shape = PROV), cex = 3, alpha = 0.75) + 
  stat_function(fun = function(x) {inv.logit(coef[1, 1] + coef[2, 1]*x + coef[3, 1]*(x^2))}, 
                size = 1.5, colour = "black") +
  scale_x_continuous(breaks=c(0, 2.5, 5, 7.5, 10)) +
  scale_y_continuous(breaks=c(0, 0.2, 0.4, 0.6, 0.8, 1)) +
  scale_colour_manual(values = cols) +
  xlab("ln seed mass (mg)") + ylab("Proportion of seeds removed") +
  theme_classic(base_size = 18) +
  theme(legend.position = "top", legend.box = "vertical", legend.margin = margin(),
        panel.border = element_rect(linetype = "solid", fill = NA),
        axis.text = element_text(colour = "black")) +
  guides(colour = guide_legend(title="Ecosystem:", order = 1), 
         shape = guide_legend(title="Provenance:", order = 2))
```


## 3. Exclude data points with missing exposure time

The pattern is similar with the result above. That is, the exclusion of these 15 data points is *not* the reason of the difference between the results of Dylewski et al. (2020) and our re-analyses (Sections 4 & 5).

Note the adding of `drop_na(TIME_D)`.

```{r}
# Transform data according to Dylewski et al. (2020)
data_C = data %>% drop_na(TIME_D) %>% 
  # "... applying a small, linear transformation to eliminate zeros and ones (Smithson & Verkuilen 2006)
  # Rescale to 0 < y < 1
  mutate(REMOVAL_C = (REMOVAL * (nrow(data) - 1) + 1/2)/nrow(data))

# Model
# "... using generalised linear mixed models (GLMMs) with a beta distribution and a restricted maximum-likelihood (REML) estimator"
model <- glmmTMB(REMOVAL_C ~ MASS_ln + I(MASS_ln^2) + (1|PUB), data = data_C, 
                 family = beta_family(link = "logit"), REML = TRUE)
summary(model); coef = summary(model)$coefficients$cond

# Graph
ggplot(data = data_C, aes(x = MASS_ln, y = REMOVAL_C)) + 
  geom_point(aes(colour = ECOSYSTEM, shape = PROV), cex = 3, alpha = 0.75) + 
  stat_function(fun = function(x) {inv.logit(coef[1, 1] + coef[2, 1]*x + coef[3, 1]*(x^2))}, 
                size = 1.5, colour = "black") +
  scale_x_continuous(breaks=c(0, 2.5, 5, 7.5, 10)) +
  scale_y_continuous(breaks=c(0, 0.2, 0.4, 0.6, 0.8, 1)) +
  scale_colour_manual(values = cols) +
  xlab("ln seed mass (mg)") + ylab("Proportion of seeds removed") +
  theme_classic(base_size = 18) +
  theme(legend.position = "top", legend.box = "vertical", legend.margin = margin(),
        panel.border = element_rect(linetype = "solid", fill = NA),
        axis.text = element_text(colour = "black")) +
  guides(colour = guide_legend(title="Ecosystem:", order = 1), 
         shape = guide_legend(title="Provenance:", order = 2))
```


## 4. A linear decline in seed survival through time

We standardised seed removal data for 24 hours of exposure to post-dispersal predators by assuming a linear decline in seed survival through time. The unimodal pattern between seed mass and seed removal vanished.

```{r}
# Standardise data
data_C = data %>% drop_na(TIME_D) %>% 
  # Linear decline
  mutate(REMOVAL_Mean = REMOVAL / TIME_D,
         REMOVAL_Mean = replace(REMOVAL_Mean, REMOVAL_Mean > 1, 1)) %>% 
  # Rescale to 0 < y < 1
  mutate(REMOVAL_C = (REMOVAL_Mean * (nrow(data) - 1) + 1/2)/nrow(data))

# Model
model <- glmmTMB(REMOVAL_C ~ MASS_ln + I(MASS_ln^2) + (1|PUB), data = data_C, 
                 family = beta_family(link = "logit"), REML = TRUE)
summary(model); coef = summary(model)$coefficients$cond

# Graph
ggplot(data = data_C, aes(x = MASS_ln, y = REMOVAL_C)) + 
  geom_point(aes(colour = ECOSYSTEM, shape = PROV), cex = 3, alpha = 0.75) + 
  stat_function(fun = function(x) {inv.logit(coef[1, 1] + coef[2, 1]*x + coef[3, 1]*(x^2))}, 
                size = 1.5, colour = "black", linetype = 5) +
  scale_x_continuous(breaks=c(0, 2.5, 5, 7.5, 10)) +
  scale_y_continuous(breaks=c(0, 0.2, 0.4, 0.6, 0.8, 1)) +
  scale_colour_manual(values = cols) +
  xlab("ln seed mass (mg)") + ylab("Proportion of seeds removed") +
  theme_classic(base_size = 18) +
  theme(legend.position = "top", legend.box = "vertical", legend.margin = margin(),
        panel.border = element_rect(linetype = "solid", fill = NA),
        axis.text = element_text(colour = "black")) +
  guides(colour = guide_legend(title="Ecosystem:", order = 1), 
         shape = guide_legend(title="Provenance:", order = 2))
```

## 5. An exponential decline in seed survival through time

We standardised seed removal data for 24 hours of exposure to post-dispersal predators by assuming an exponential decline in seed survival through time. The unimodal pattern remained significant but became very shallow.

```{r}
# Standardise data
data_C = data %>% drop_na(TIME_D) %>% 
  # Exponential decline
  mutate(SURVIVAl = 1 - REMOVAL,
         REMOVAL_Exp = 1 - SURVIVAl^(1/TIME_D)) %>% 
  # Rescale to 0 < y < 1
  mutate(REMOVAL_C = (REMOVAL_Exp * (nrow(data) - 1) + 1/2)/nrow(data))

# Model
model <- glmmTMB(REMOVAL_C ~ MASS_ln + I(MASS_ln^2) + (1|PUB), data = data_C, 
                 family = beta_family(link = "logit"), REML = TRUE)
summary(model); coef = summary(model)$coefficients$cond

# Graph
ggplot(data = data_C, aes(x = MASS_ln, y = REMOVAL_C)) + 
  geom_point(aes(colour = ECOSYSTEM, shape = PROV), cex = 3, alpha = 0.75) + 
  stat_function(fun = function(x) {inv.logit(coef[1, 1] + coef[2, 1]*x + coef[3, 1]*(x^2))}, 
                size = 1.5, colour = "black") +
  scale_x_continuous(breaks=c(0, 2.5, 5, 7.5, 10)) +
  scale_y_continuous(breaks=c(0, 0.2, 0.4, 0.6, 0.8, 1)) +
  scale_colour_manual(values = cols) +
  xlab("ln seed mass (mg)") + ylab("Proportion of seeds removed") +
  theme_classic(base_size = 18) +
  theme(legend.position = "top", legend.box = "vertical", legend.margin = margin(),
        panel.border = element_rect(linetype = "solid", fill = NA),
        axis.text = element_text(colour = "black")) +
  guides(colour = guide_legend(title="Ecosystem:", order = 1), 
         shape = guide_legend(title="Provenance:", order = 2))
```